### 2021-05-27 written/eLifeRevision/Response.tex

We thank the reviewers for emphasizing the importance of this methodological comparison.  In Section 3.3, we provide a direct comparison of EPI to alternative simulation-based inference techniques SMC-ABC and SNPE by inferring RNN connectivities that exhibit stable amplification.  These comparisons evaluate both wall time (Fig. 2A) and simulation count (Fig. 2B), and we explain how each algorithm was run in its preferred hardware setting in Section 5.3.4.

In this analysis, we demonstrate the improved scalability of deep inference techniques (EPI and SNPE) with respect to the state-of-the-art approximate bayesian computation technique (SMC-ABC).
While controlling for architecture size (Fig. RNN1), we push the limits of SNPE through targeted hyperparameter modifications (Fig. RNN3), and show that EPI scales to higher dimensional RNN connectivities producing stable amplification than SNPE.
Furthermore, we emphasize that SNPE does not constrain the properties of the inferred parameters; many connectivities inferred by SNPE result in unstable or nonamplified models (Fig. 2C, Fig. RNN2).

\textbf{3. The extent to which this method is generally/straightforwardly applicable was in doubt. It seemed as though a significant amount of computation was required to do inference on one specified property and that the computation would need to be run afresh to query a new property. The methodology in the cosubmission (Macke) made clear that computation required for successive inferences is 'amortized' during training on random parameters. Moreover, EPI seemed less flexible than the cosubmission's approach in that it required a differentiable loss function. The complementarity and advantages of this approach as opposed to the cosubmission's approach are therefore unclear. }

We respectfully disagree with this characterization, but we take responsibility here: we certainly needed to improve our exposition and comparisons to clarify the general applicability of EPI.  We have done so, and what results is a meaningful comparison with the cosubmission (SNPE, Macke et al, now published) and other methods.  Of course, there is a tradeoff (no free lunch, as usual), and these new analyses clarify when EPI is preferable to SNPE, and vice versa.

@@

We thank the reviewer for their positive comments.

\textbf{Major points }

\textbf{1. Training deep networks is hard. Indeed the authors devote a substantial amount of the manuscript to techniques for training them, and note that different hyperparameters were necessary for each of the different studies. Can the authors be confident that they have found the network which gives maximum entropy or close to it? If so, how. If not, how does that affect the conclusions? }

We agree with the reviewer that hyperparameter sensitivity and global optima are important considerations of any optimization algorithm using deep neural networks.  To draw a parallel, training deep networks for visual processing used to be considered infeasible, but became easier through iterative improvements in architectural and hyperparameter choices that spread across the field.  
Similarly, training deep networks via EPI to learn parameter distributions became easier throughout this research project as we learned through trial and error what works well.  
In fact, there has been extraordinary progress in the field of deep probability distributions (specifically normalizing flows), that have allowed EPI to converge regularly while capturing complex structure (e.g. Dinh et al. 2017 and Kingma et al. 2018).  
This manuscript's explanation of hyperparameter choices and the extensive set of examples in the online code will serve as valuable guidelines for future research using this method.
Every figure of this paper is reproducible with the jupyter notebooks, and there are several tutorials for understanding the most consequential hyperparameters: the augmented lagrangian constant and deep probability distribution architecture.

### 2021-05-27 written/eLifeRevision/Bittner_eLife2020.tex
update algo and pass through paper.

\begin{figure}
\begin{center}
\includegraphics[scale=.8]{figures/fig2/fig2.pdf}
\end{center}
\caption{ 
\textbf{A}. Wall time of EPI (blue), SNPE (orange), and SMC-ABC (green) to converge on RNN connectivities producing stable amplification.  
Each dot shows convergence time for an individual random seed.
For reference, the mean wall time for EPI to achieve its full constraint convergence (means and variances) is shown (blue line).
\textbf{B}. Simulation count of each algorithm to achieve convergence. Same conventions as A.
\textbf{C}. The predictive distributions of connectivities inferred by EPI (blue), SNPE (orange), and SMC-ABC (green), with reference to $\mathbf{x}_0 = \bm{\mu}$ (gray star).
\textbf{D}. Simulations of networks inferred by each method ($\tau=100ms$).  Each trace (15 per algorithm) corresponds to simulation of one $z$.  (Below) Ratio of obtained samples producing stable amplification, monotonic decay, and instability.
}
\label{fig:RNN}
\end{figure}

@@

\q{v1_ep_text1}{We quantify levels of E-population variability by studying two emergent properties}
\begin{equation}\label{eq:EP_V1}
\begin{split}
\mathcal{X}(5\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 5\text{Hz}~~~~~~~\mathcal{X}(10\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~ 10\text{Hz}  \\ 
 \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2~~~~~~~~~~~~~~~~~~~~~~~ \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  1 \text{Hz}^2,
\end{split}
\end{equation}
\q{v1_ep_text2}{where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
In the following analyses, we select 1Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.}

@@

\subsection{Emergent property inference (EPI)}\label{methods_EPI}
Solving inverse problems is an important part of theoretical neuroscience, since we must understand how neural circuit models and their parameter choices can produce computations of varying levels of complexity.
While much machine learning research has focused on how to find latent structure in large-scale neural datasets, less has focused on inverting theoretical circuit models with respect to their computational properties.
Here, we introduce a novel method for statistical inference, which uses deep networks to learn parameter distributions constrained to produce emergent properties of computation.

@@

In this work, emergent properties $\mathcal{X}$ are defined through the choice of emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (which is a vector of one or more statistics), and its means $\bm{\mu}$, and variances $\bm{\sigma}^2$:
 \begin{equation} \label{eq:methods_ep_general}
\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
\end{equation}
In general, an emergent property may be a collection of first-, second-, or higher-order moments of a group of statistics, but this study focuses on the case written in Equation \ref{eq:methods_ep_general}.
In the STG example, intermediate hub frequency is defined by mean and variance constraints on the statistic of hub neuron frequency $\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})$ (Equations \ref{eq:EP_STG1} and \ref{eq:EP_STG2}).
Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$ over the EPI distribution of parameters ($\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$) and the data produced by those parameters ($\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$), where the learned parameter distribution $q_{\bm{\theta}}(z)$ is determined by deep network weights and biases $\bm{\theta}$.

In EPI, deep probability distributions are optimized to approximate the inferred distribution.
 In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ (we choose an isotropic gaussian) is mapped deterministically via a sequence of deep neural network layers ($g_1$, .. $g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
\begin{equation}
\label{eq:deep_transform}
\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
\end{equation}
Such deep probability distributions embed the inferred distribution in a deep network.
Once optimized, this deep network representation of a distribution has remarkably useful properties: fast sampling and probability evaluations.
Importantly, fast probability evaluations confer fast gradient and Hessian calculations as well.

Given this choice of circuit model and emergent property $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$ that produces the emergent property $\mathcal{X}$:
\begin{equation} \label{eq:opt}
\begin{split}
q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) = q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
 &  \text{s.t.  } \mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2, \\
 \end{split}
\end{equation} 
where $H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}}\left[- \log q_{\bm{\theta}}(z) \right]$ is entropy.
By maximizing the entropy of the inferred distribution $q_{\bm{\theta}}$, we select the most random distribution in family $\mathcal{Q}$ that satisfies the constraints of the emergent property.
Because entropy is maximized in Equation \ref{eq:opt}, EPI is equivalent to bayesian variational inference (see Section \ref{methods_VI}).
\q{AL_standard}{
To run this constrained optimization, we use an augmented lagrangian objective, which is the standard approach for constrained optimization \cite{bertsekas2014constrained}, and the approach taken to fit Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.}
This procedure is detailed in Section \ref{methods_AL_opt} and the pseudocode in Algorithm \ref{alg:epi}.

@@

The method EPI differs from SNPE in some key ways.
SNPE belongs to a ``sequential" class of recently developed simulation-based inference methods in which two neural networks are used for posterior inference.
This first neural network is a deep probability distribution (normalizing flow) used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL) \cite{papamakarios2019sequential}).
A recent advance uses an unconstrained neural network to estimate the likelihood ratio (sequential neural ratio estimation (SNRE) \cite{hermans2020likelihood}).
In SNL and SNRE, MCMC sampling techniques are used to obtain samples from the approximated posterior.
This contrasts with EPI and SNPE, which use deep probability distributions to model parameters, which facilitates immediate measurements of sample probability, gradient, or Hessian for system analysis.
The second neural network in this sequential class of methods is the amortizer.  This unconstrained deep network maps data $\mathbf{x}$ (or statistics $f(\mathbf{x}; \mathbf{z})$ or model parameters $\mathbf{z}$ to the weights and biases of the first neural network.
These methods are optimized on a conditional density (or ratio) estimation objective.
The data used to optimize this objective are generated via an adaptive procedure, in which training data pairs ($\mathbf{x}_i$, $\mathbf{z}_i$) become sequentially closer to the true data and posterior.

\q{amortization}{The approximating fidelity of the deep probability distribution in sequential approaches is optimized to generalize across the training distribution of the conditioning variable.
This generalization property of the sequential methods can reduce the accuracy at the singular posterior of interest.
Whereas in EPI, the entire expressivity of the deep probability distribution is dedicated to learning a single distribution as well as possible.
Amortization is not possible in EPI, since EPI learns an exponential family distribution parameterized by its mean (in contrast to its natural parameter, see Section \ref{methods_ME_EF}).}
The well-known inverse mapping problem of exponential families \cite{wainwright2008graphical} prohibits an amortization based approach in EPI, since emergent properties are defined via the mean parameter of the exponential family distribution.
However, we have shown that the same two-network architecture of the sequential simulation-based inference methods can be used for amortized inference in intractable exponential family posteriors when using their natural parameterization \cite{bittner2019approximating}.

@@

In this manuscript, emergent properties are defined by statistics $f(\mathbf{x}; \mathbf{z})$ having a fixed mean $\bm{\mu}$ and variance $\bm{\sigma}^2$ as in Equation \ref{eq:methods_ep_general}.
The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$:
 \begin{equation}
\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right].
\end{equation}
As a general maximum entropy distribution (Equation \ref{eq:max_ent}), the sufficient statistics vector contains both first and second order moments of $f(\mathbf{x}; \mathbf{z})$
\begin{equation} \label{eq:moments}
T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2 \end{bmatrix},
\end{equation}
which are constrained to the chosen means and variances
\begin{equation} \label{eq:mu_opt}
\bm{\mu}_{\text{opt}} = \begin{bmatrix} \bm{\mu} \\ \bm{\sigma}^2 \end{bmatrix}.
\end{equation}

Thus, $\bm{\mu}_{\text{opt}}$ is used to denote the mean parameter of the maximum entropy distribution defined by the emergent property (all constraints), while $\bm{\mu}$ is only the mean of $f(\mathbf{x}; \mathbf{z})$.
The subscript ``$\text{opt}$" of $\bm{\mu}_{\text{opt}}$ is chosen since it contains all of the contraint values to which the optimization algorithm must adhere.

\subsubsection{Augmented lagrangian optimization}\label{methods_AL_opt}
To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained maximum entropy optimization is executed using the augmented lagrangian method.  
The following objective is minimized:
\begin{equation} \label{eq:AL}
L(\bm{\theta}; \bm{\eta}_{\text{opt}}, c) = -H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top R(\bm{\theta}) + \frac{c}{2}||R(\bm{\theta})||^2
\end{equation}
where there are average constraint violations 
\begin{equation}
R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right],
\end{equation}
 $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the lagrange multipliers where $m$ is the number of total constraints
\begin{equation}
m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|,
\end{equation}
and $c$ is the penalty coefficient. 
The mean parameter $\bm{\mu}_{\text{opt}}$ and sufficient statistics $T(\mathbf{x}; \mathbf{z})$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
Specifically, $T(\mathbf{x}; \mathbf{z})$ is a concatenation of the first and second moments (Equation \ref{eq:moments}) and $\bm{\mu}_{\text{opt}}$ is a concatenation of $\bm{\mu}$ and $\bm{\sigma}^2$ (Equation \ref{eq:mu_opt}).
(Although, note that this algorithm is written for general $T(\mathbf{x}; \mathbf{z})$ and $\bm{\mu}$ to satisfy the more general class of emergent properties.)
The lagrange multipliers $\bm{\eta}_{\text{opt}}$ are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
Weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.

The gradient with respect to entropy $H(q_{\bm{\theta}}(\mathbf{z}))$ can be expressed using the reparameterization trick as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0(\mathbf{z}_0$):
\begin{equation}
H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
\end{equation}
Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average with respect to the base distribution $\mathbf{z}_0$:
\begin{equation}
\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
\end{equation}

>>>>

\begin{figure}
\begin{center}
\includegraphics[scale=.8]{figures/fig2/fig2.pdf}
\end{center}
\caption{ 
\textbf{A}. Wall time of EPI (blue), SNPE (orange), and SMC-ABC (green) to converge on RNN connectivities producing stable amplification.  
Each dot shows convergence time for an individual random seed.
For reference, the mean wall time for EPI to achieve its full constraint convergence (means and variances) is shown (blue line).
\textbf{B}. Simulation count of each algorithm to achieve convergence. Same conventions as A.
\textbf{C}. The predictive distributions of connectivities inferred by EPI (blue), SNPE (orange), and SMC-ABC (green), with reference to $\mathbf{x}_0 = \bm{\mu}$ (gray star).
\textbf{D}. Simulations of networks inferred by each method ($\tau=100ms$).  Each trace (15 per algorithm) corresponds to simulation of one $z$.  (Below) Ratio of obtained samples producing stable amplification, stable monotonic decay, and instability.
}
\label{fig:RNN}
\end{figure}

@@

\q{v1_ep_text1}{We quantify levels of E-population variability by studying two emergent properties}
\begin{equation}\label{eq:EP_V1}
\begin{split}
\mathcal{X}(5\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}, 
\mathbf{x}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 5\text{Hz}~~~~~~~\mathcal{X}(10\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}, 
\mathbf{x}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~ 10\text{Hz}  \\ 
 \text{Var}_{\mathbf{z}, 
\mathbf{x}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2~~~~~~~~~~~~~~~~~~~~~~~ \text{Var}_{\mathbf{z}, 
\mathbf{x}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  1 \text{Hz}^2,
\end{split}
\end{equation}
\q{v1_ep_text2}{where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
In the following analyses, we select 1Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.}

@@

\subsection{Emergent property inference (EPI)}\label{methods_EPI}
Solving inverse problems is an important part of theoretical neuroscience, since we must understand how neural circuit models and their parameter choices produce computations.
Recently, research on machine learning methodology for neuroscience has focused on finding latent structure in large-scale neural datasets, while research in theoretical neuroscience generally focuses on developing precise neural circuit models that can produce computations of interest.
By quantifying computation into an \emph{emergent property} through statistics of the emergent activity of neural circuit models, we can adapt the modern technique of deep probabilistic inference towards solving inverse problems in theoretical neuroscience.
Here, we introduce a novel method for statistical inference, which uses deep networks to learn parameter distributions constrained to produce emergent properties of computation.

@@

In this work, emergent properties $\mathcal{X}$ are defined through the choice of emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (which is a vector of one or more statistics), and its means $\bm{\mu}$, and variances $\bm{\sigma}^2$:
 \begin{equation} \label{eq:methods_ep_general}
\mathcal{X}:~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu},~~ \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
\end{equation}
In general, an emergent property may be a collection of first-, second-, or higher-order moments of a group of statistics, but this study focuses on the case written in Equation \ref{eq:methods_ep_general}.
In the STG example, intermediate hub frequency is defined by mean and variance constraints on the statistic of hub neuron frequency $\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})$ (Equations \ref{eq:EP_STG1} and \ref{eq:EP_STG2}).
Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$ over the EPI distribution of parameters ($\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$) and the data produced by those parameters ($\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$), where the inferred parameter distribution $q_{\bm{\theta}}(\mathbf{z})$ itself is parameterized by deep network weights and biases $\bm{\theta}$.

In EPI, a deep probability distribution $q_{\bm{\theta}}(\mathbf{z})$ is optimized to approximate the parameter distribution producing the emergent property $\mathcal{X}$.
In contrast to simpler classes of distributions like the gaussian or mixture of gaussians, deep probability distributions are far more flexible and capable of fitting rich structure \cite{rezende2015variational, papamakarios2019normalizing}.
 In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ (we choose an isotropic gaussian) is mapped deterministically via a sequence of deep neural network layers ($g_1$, .. $g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
\begin{equation}
\label{eq:deep_transform}
\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
\end{equation}
Such deep probability distributions embed the inferred distribution in a deep network.
Once optimized, this deep network representation of a distribution has remarkably useful properties: fast sampling and probability evaluations.
Importantly, fast probability evaluations confer fast gradient and Hessian calculations as well.

Given this choice of circuit model and emergent property $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$ that produces the emergent property $\mathcal{X}$:
\begin{equation} \label{eq:opt}
\begin{split}
q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) = q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
 &  \text{s.t.  } \mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2, \\
 \end{split}
\end{equation} 
where $H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}}\left[- \log q_{\bm{\theta}}(z) \right]$ is entropy.
By maximizing the entropy of the inferred distribution $q_{\bm{\theta}}$, we select the most random distribution in family $\mathcal{Q}$ that satisfies the constraints of the emergent property.
Since entropy is maximized in Equation \ref{eq:opt}, EPI is equivalent to bayesian variational inference (see Section \ref{methods_VI}), which is why we specify the inferred distribution of EPI as conditioned upon emergent property $\mathcal{X}$ with the notation $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
\q{AL_standard}{
To run this constrained optimization, we use an augmented lagrangian objective, which is the standard approach for constrained optimization \cite{bertsekas2014constrained}, and the approach taken to fit Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.}
This procedure is detailed in Section \ref{methods_AL_opt} and the pseudocode in Algorithm \ref{alg:epi}.

@@

The method EPI differs from SNPE in some key ways.
SNPE belongs to a ``sequential" class of recently developed simulation-based inference methods in which two neural networks are used for posterior inference.
This first neural network is a deep probability distribution (normalizing flow) used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL) \cite{papamakarios2019sequential}).
A recent approach uses an unconstrained neural network to estimate the likelihood ratio (sequential neural ratio estimation (SNRE) \cite{hermans2020likelihood}).
In SNL and SNRE, MCMC sampling techniques are used to obtain samples from the approximated posterior.
This contrasts with EPI and SNPE, which use deep probability distributions to model parameters, which facilitates immediate measurements of sample probability, gradient, or Hessian for system analysis.
The second neural network in this sequential class of methods is the amortizer.  This unconstrained deep network maps data $\mathbf{x}$ (or statistics $f(\mathbf{x}; \mathbf{z})$ or model parameters $\mathbf{z}$) to the weights and biases of the first neural network.
These methods are optimized on a conditional density (or ratio) estimation objective.
The data used to optimize this objective are generated via an adaptive procedure, in which training data pairs ($\mathbf{x}_i$, $\mathbf{z}_i$) become sequentially closer to the true data and posterior.

\q{amortization}{The approximating fidelity of the deep probability distribution in sequential approaches is optimized to generalize across the training distribution of the conditioning variable.
This generalization property of the sequential methods can reduce the accuracy at the singular posterior of interest.
Whereas in EPI, the entire expressivity of the deep probability distribution is dedicated to learning a single distribution as well as possible.
The well-known inverse mapping problem of exponential families \cite{wainwright2008graphical} prohibits an amortization-based approach in EPI, since EPI learns an exponential family distribution parameterized by its mean (in contrast to its natural parameter, see Section \ref{methods_ME_EF}).}
However, we have shown that the same two-network architecture of the sequential simulation-based inference methods can be used for amortized inference in intractable exponential family posteriors when using their natural parameterization \cite{bittner2019approximating}.

@@

In this manuscript, emergent properties are defined by statistics $f(\mathbf{x}; \mathbf{z})$ having a fixed mean $\bm{\mu}$ and variance $\bm{\sigma}^2$ as in Equation \ref{eq:methods_ep_general}.
The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$:
 \begin{equation}
\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right].
\end{equation}
As a general maximum entropy distribution (Equation \ref{eq:max_ent}), the sufficient statistics vector contains both first and second order moments of $f(\mathbf{x}; \mathbf{z})$
\begin{equation} \label{eq:moments}
T(\mathbf{z}) = \begin{bmatrix} \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ f(\mathbf{x}; \mathbf{z}) \right] \\ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2\right] \end{bmatrix},
\end{equation}
which are constrained to the chosen means and variances
\begin{equation} \label{eq:mu_opt}
\bm{\mu}_{\text{opt}} = \begin{bmatrix} \bm{\mu} \\ \bm{\sigma}^2 \end{bmatrix}.
\end{equation}

Thus, $\bm{\mu}_{\text{opt}}$ is used to denote the mean parameter of the maximum entropy distribution defined by the emergent property (all constraints), while $\bm{\mu}$ is only the mean of $f(\mathbf{x}; \mathbf{z})$.
The subscript ``$\text{opt}$" of $\bm{\mu}_{\text{opt}}$ is chosen since it contains all of the constraint values to which the EPI optimization algorithm must adhere.

\subsubsection{Augmented lagrangian optimization}\label{methods_AL_opt}
To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained maximum entropy optimization is executed using the augmented lagrangian method.  
The following objective is minimized:
\begin{equation} \label{eq:AL}
L(\bm{\theta}; \bm{\eta}_{\text{opt}}, c) = -H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top R(\bm{\theta}) + \frac{c}{2}||R(\bm{\theta})||^2
\end{equation}
where there are average constraint violations 
\begin{equation}
R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})}\left[T(\mathbf{z}) - \bm{\mu}_{\text{opt}} \right],
\end{equation}
 $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the lagrange multipliers where $m$ is the number of total constraints
\begin{equation}
m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|,
\end{equation}
and $c$ is the penalty coefficient. 
The mean parameter $\bm{\mu}_{\text{opt}}$ and sufficient statistics $T(\mathbf{z})$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
Specifically, $T(\mathbf{z})$ is a concatenation of the first and second moments (Equation \ref{eq:moments}) and $\bm{\mu}_{\text{opt}}$ is a concatenation of their constraints $\bm{\mu}$ and $\bm{\sigma}^2$ (Equation \ref{eq:mu_opt}).
(Although, note that this algorithm is written for general $T(\mathbf{z})$ and $\bm{\mu}_{\text{opt}}$ to satisfy the more general class of emergent properties.)
The lagrange multipliers $\bm{\eta}_{\text{opt}}$ are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
Weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.

The gradient with respect to entropy $H(q_{\bm{\theta}}(\mathbf{z}))$ can be expressed using the reparameterization trick as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0(\mathbf{z}_0$):
\begin{equation}
H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
\end{equation}
Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients with respect to the base distribution $\mathbf{z}_0$:
\begin{equation}
\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
\end{equation}
The gradients of the log density of the deep probability distribution are tractable through the use of normalizing flows (see Section \ref{methods_NF}).

The full EPI optimization algorithm is detailed in Algorithm \ref{alg:epi}. The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic gradient descent (SGD) iterations. 
A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
The p-value of $\mathbb{E}[||R(\bm{\theta}_{k+1})||] > \gamma \mathbb{E} \left[||R(\bm{\theta}_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{z}^{(i)}) - \bm{\mu}_{\text{opt}})$ given a batch size $n$ and $\mathbf{z}^{(i)} \sim q_{\bm{\theta}}(\mathbf{z})$.
Throughout the study, $\gamma = 0.25$, while $\beta$ was chosen to be either 2 or 4.  The batch size of EPI also varied according to application.
